---
title: "Data Science and Machine Learning"
subtitle: "COM747"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Course Overview

[Course overview text]

## Prerequisites

[List prerequisites]

## Required Packages

```{r packages, eval=FALSE}
# Package installation code
```

# Course Materials

## Week 1

### Lectures

-   [Lecture 1](week1/Week%201%20-%20Introduction%20to%20Data%20Science.rmd)

### Practicals

-   [Practical 1](week1/practical1.html)

## Week 2

### Lectures

-   [Lecture 2](week2/lecture2.html)

### Practicals

-   [Practical 2](week2/practical2.html)

## Week 3

### Lectures

-   [Lecture 3](week3/lecture3.html)

### Practicals

-   [Practical 3](week3/practical3.html)

## Week 4

### Lectures

-   [Lecture 4](week4/lecture4.html)

### Practicals

-   [Practical 4](week4/practical4.html)

## Week 5

### Lectures

-   [Lecture 5](week5/lecture5.html)

### Practicals

-   [Practical 5](week5/practical5.html)

## Week 6

### Lectures

-   [Lecture 6](week6/lecture6.html)

### Practicals

-   [Practical 6](week6/practical6.html)

# Assessment

[Assessment information]

# Additional Resources

[List of resources]

# Getting Help

[Support information]
