# Week 2 - Exercise 3

Download rundata.csv from BBL and open in Excel

-   R Tutorial – insert the following code

Import rundata.csv into R Studio View(rundata)

From Text Base

```{r}
rundata$FirstRun
```

```{r}
mean(rundata$FirstRun)
```

```{r}
sd(rundata$FirstRun)
```

```{r}
summary(rundata$FirstRun)
```

```{r}
hist(rundata$FirstRun)
```

## Spend time understanding the other parameters of the hist() function

```{r}
?hist
hist(rundata$FirstRun, ylab="y-axis title", xlab="x-axis title", main="main title")
dens <- density(rundata$FirstRun)
plot(dens)
```

## Computing the probability of running a distance of 20 to 25km

```{r}
convertToFunc <- with(dens, splinefun(x,y))
integrate(convertToFunc, lower=20, upper=25)
```

## Normality Testing

### Shapiro-Wilk test of normality.

```{r}
shapiro.test(rundata$FirstRun)
```

### Hypothesis test

$H_0 = \text{Null hypotheseis}:\text{Data is normally distributed}$

$H_1 = \text{Alternative hypotheseis}:\text{Data is NOT normally distributed}$

```{r}
hist(rundata$FirstRun)
hist(rundata$SecondRun)
```

```{r}
plot(density(x = rundata$FirstRun))
plot(density(x = rundata$SecondRun))
```

Write this code to create overlapping density functions and try to interpret the
results.

```{r}
plot(density(rundata$FirstRun), col="red")
lines(density(rundata$SecondRun), col="green")
legend("bottom", legend = c("First run","Second run"), fill=c("red","green"))
boxplot(rundata)
?boxplot
```

```{r}
plot(rundata$FirstRun, rundata$SecondRun)
```
