# Week 2 - Exercise 2

# **1. The data frame() function:**

Data frames are like spreadsheets and can be created or imported using any of
the following functions:

```{r}
data.frame()
read.table()
read.csv()
```

Create your own data frame:

```{r}
**x <- data.frame(names= c(“ben”, “joe”, “steve”), ages= c(31, 24, 45))**
```

Now use the following functions to determine the number of columns and rows:

```{r}
**nrow(x)
ncol(x)**
```

Now use the structure function – to learn more type **`?str`**

```{r}
**str(x)**
```

Now you can access any column/variable in a data frame using the dollar sign.
Try the following:

```{r}
**x$names**
```

Note ‘x’ is the name of the data frame and ‘names’ is the name of the column.

Type in the following to see a list of free data frames in R:

```{r}
**data()**
```

Type in the following

```{r}
**x <-mtcars**
```

A new dataframe **x** should appear in your environment. You can read about it
by typing in the following:

```{r}
**?mtcars**
```

Now have a go at analyzing this data.

Boxplot miles per galloon and create a histogram:

```{r}
**boxplot(mtcars$mpg)
hist(mtcars$mpg)**
```

Boxplot miles per galloon for manual and automatic cars.

First method – sub-setting approach:

```{r}
**par(mfrow=c(2,1))
boxplot(mtcars$mpg[mtcars$am==1], main=”auto cars”)
boxplot(mtcars$mpg[mtcars$am==0], main=”manual cars”)**
```

Second method:

```{r}
**boxplot(mtcars$mpg~mtcars$am)**
```

# **2. Ggplot visualisations**

Install ggplot by typing in the following code:

```{r}
**install.packages("ggplot2")**
```

Once installed, you can activate the library using the following code:

```{r}
library(ggplot2)
```

Now try this code (same as above only in ggplot):

```{r}
p <- ggplot(mtcars, aes(factor(am), mpg))
p + geom_boxplot()
```

Now add a jitter to show density:

```{r}
**p <- ggplot(mtcars, aes(factor(am), mpg))
p + geom_boxplot() + geom_jitter(width = 0.2)**
```

Add a third dimension – engine size (vs):

```{r}
p <- ggplot(mtcars, aes(factor(am), mpg))
p + geom_boxplot(aes(colour = factor(vs)))
```
