---
title: "Week 2 - Exercise 1"
output: rmarkdown::github_document
---

# Week 2 - Exercise 1

Continuing on from last week, try the following code snippets:

# 1. The apply() function:

Read about the apply function by typing in

```{r}
?apply
```

*The structure of the apply function is as follows:*

```{r}
apply(m, dimcode, f, fargs)
```

Now try the apply function with the mean function:

```{r}
x <- matrix(1:20, nrow = 10, ncol = 2)
apply(x, 2, mean)
```

Now try the following code and try to understand each statement:

```{r}
x <- matrix(1:20, nrow = 10, ncol = 2)
rowSums <- apply(x, 1, sum)
rowMeans <- apply(x, 1, mean)
colSums <- apply(x, 2, sum)
colMeans <- apply(x, 2, mean)
```

Now try this code using the quantile function with the probs argument. Note that
the probs argument is of the quantile function as opposed to the apply function.

```{r}
apply(x, 1, quantile, probs = c(0.25, 0.75))
```

# 2. Factors

Factors represent nominal/categorical data.

Try the following code:

```{r}
?factor
x <- factor( c("yes", "yes", "no", "yes", "no") )
x
```

The print out should be:

```{r}
yes yes no yes no 
Levels: no yes
```

Note: levels is a statistical term meaning categories.

# 3. The table function:

Now try the table() function in R. This function takes a factor variable and
provides the number of occurrences of each category/level in that variable.

```{r}
x <- factor( c("yes", "yes", "no", "yes", "no") )
table(x)
```

The print out should be:

```{r}
no yes
23
```

# 4. The tapply() function:

The tapply() function is a mixture of the table and apply functions. The tapply
function runs a function across two variables, for example you can run the mean
function to calculate the average age of people in each subgroup (by
affiliation). Try the following code:

```{r}
ages <- c(25,26,37,56,42,25)
affill <- factor(c("U", "R", "R", "D", "D", "U")) 
tapply(ages, affill , mean)
tapply(ages, affill , range)
```

# 5. Handling missing values:

Try the following code:

```{r}
x <- c(25,26,NA,56,42,25)
is.na(x)
is.nan(x)
```

The function is.na(x) is used to test for NA and returns a logical vector. The
function is.nan(x) is used to test for NaN and returns a logical vector.

Try the following code:

```{r}
x <- c(1, 2, NA, 10, 3)
is.na(x)
# Print out should be: [1] FALSE FALSE TRUE FALSE FALSE*

is.nan(x)
# *Print out should be: [1] FALSE FALSE FALSE FALSE FALSE*

x <- c(1, 2, NaN, NA, 4)
is.na(x)
# *Print out should be: [1] FALSE FALSE TRUE TRUE FALSE*

is.nan(x)
# *Print out should be: [1] FALSE FALSE TRUE FALSE FALSE*
```

# 6. Data cleansing of missing values:

Try the following two approaches for cleansing missing values. The first
approach sets all missing values to 0 and the second just inputs the mean
average.

```{r}
x <- c(1, 2, NA, 10, 3)
x[is.na(x)] <- 0
x <- c(1, 2, NA, 10, 3)
x[is.na(x)] <- mean(x)
```

# 7. The mean function with missing values:

If you try the following code, you will notice that it does not work as
expected:

```{r}
x <- c(1, 2, NA, 10, 3)
mean(x)
```

To work as expected you need to use the na.rm argument to remove NA values:

```{r}
mean(x, na.rm=TRUE)
```
