# Week 2 - Exercise 4

# **Other probability distributions**

The normal distribution models most variables but not all variables are
continuous numbers and not all continuous numbers have a bell shape, hence you
may need to use a different distribution to model and approximate some
variables.

## **Binomial distributions**

This distribution models events that have binary outcomes (e.g. pass or fail,
yes or no).

Imagine you had 20 multiple choice questions. Each question has 4 answers (only
one is correct), hence the outcome for each question is either correct or
incorrect. The probability of getting one question correct at random guess is ¼
(0.25 or 25%).

## **If you answer the questions randomly, what is the probability of getting precisely 50% of the questions correct:**

```{r}
dbinom(10, size=20, prob=0.25)
```

## **What is the probability of getting 50% or less questions correct?**

This is the cumulative probability function:

```{r}
pbinom(10, size=20, prob=0.25)
```

# **Poisson distribution**

You can use google to see what a Poisson distribution looks like. It models the
probability of events that occur per a time interval – for example:

## **On average, per minute, if there are 7 people in a coffee shop queue, what is the probability that there are 6 or less people in this queue?**

```{r}
ppois(6, lambda=7)
```

## **The probability of having 6 or more people in the queue is the inverse of this:**

```{r}
1 - ppois(6, lambda=7)
```

or you can write this code:

```{r}
ppois(6, lambda=7, lower=FALSE)
```

# **Exponential distribution**

**Imagine you deliver food for fast food outlet, if your food delivery times fit
an exponential distribution, and if your mean delivery time is 10 minutes, what
is the probability of delivering food in less than 5 minutes?**

## The rate of delivery in the distribution is 1 delivery every 10 minutes, hence 1/10.

```{r}
pexp(5, rate=1/10)
```

Feel free to complete similar R code demonstrations for different distributions:
<https://www.r-tutor.com/elementary-statistics/probability-distributions/chi-squared-distribution>

**Reference:**

**This tutorial is adapted from:**

<https://www.r-tutor.com/elementary-statistics/probability-distributions>
