# Week 3 - Exercise 1

# **Law of large numbers exercise:**

The law of large numbers could be described as the fact that as the sample
size/dataset gets larger, its sample mean and sample SD moves closer to the
population mean and SD. In other words, larger sample sizes are typically more
representative of the population.

## **Write the following code into R studio:**

```{r}
population <- rnorm(1000, 52, 40)
samplemeans <- vector(length=100)
sampleSDs <- vector(length=100)
for(i in 1:100){
    y <- sample(population, i*10)
    samplemeans[i] <- mean(y)
    sampleSDs[i] <- sd(y)
}
```

This code demonstrates the concept of sampling distribution and the effect of
increasing sample size on sample statistics. Here's what each part does:

1.  `population <- rnorm(1000, 52, 40)` creates a population of 1,000 values
    drawn from a normal distribution with mean 52 and standard deviation 40.

2.  `samplemeans <- vector(length=100)` and `sampleSDs <- vector(length=100)`
    create empty vectors to store the results.

3.  The `for` loop runs 100 iterations. In each iteration:

    -   `y <- sample(population, i*10)` takes a random sample from the
        population. The sample size increases by 10 in each iteration (10 in
        first iteration, 20 in second, etc., up to 1000).
    -   `samplemeans[i] <- mean(y)` calculates the mean of the current sample
        and stores it.
    -   `sampleSDs[i] <- sd(y)` calculates the standard deviation of the current
        sample and stores it.

The purpose of this code is to demonstrate how sample statistics (mean and
standard deviation) become more stable and closer to the true population
parameters as sample size increases. This illustrates the law of large numbers -
as sample size grows, the sample mean converges to the population mean (52) and
the sample variability decreases.

If you plotted `samplemeans` against the sample sizes, you'd likely see the
means stabilizing around 52 as the sample size increases. Similarly, `sampleSDs`
would tend to converge toward 40 with larger samples.

-   **Plot the following, and notice how the sample mean is more akin to the
    population mean as the sample size increases**

```{r}
plot(samplemeans)
abline(h=52)
```

-   **Plot the following and notice how the sample SD is more akin to the
    population SD as the sample size increases**

```{r}
plot(sampleSDs)
abline(h=40)
```

-   **Time permitting, have some fun playing with the simulator (code) and
    change the parameters etc., for example the population statistics. And note
    down your observations.**
